---
title: "HP2: Board Games"
author: "Brian Hsu"
format: html  
editor: visual
---

## Introduction

### **Objective**

Explore the patterns in board game design, ratings, and user preferences using the BoardGameGeek dataset. The study focuses on game specifications, player count, and publication trends, aiming to derive insights into factors contributing to game popularity.

### **Methods**

The dataset was analyzed using R. Techniques included data cleaning, aggregation, and visualization to explore trends in game characteristics and ratings

## Data Dictionary

#### `ratings.csv`

| variable      | class     | description                                                                                    |
|:------------------|:------------------|:---------------------------------|
| num           | double    | A unique identifier for each game within the dataset.                                          |
| id            | double    | The BoardGameGeek ID for the game, used for linking to other datasets or resources.            |
| name          | character | The official name of the board game, as listed on BoardGameGeek.                               |
| year          | double    | The year the game was first published or made available to the public.                         |
| rank          | double    | The global rank of the game based on user ratings, where lower ranks indicate higher ratings.  |
| average       | double    | The average user rating for the game, calculated across all submitted ratings.                 |
| bayes_average | double    | A Bayesian-adjusted average rating, incorporating a prior to reduce the influence of outliers. |
| users_rated   | double    | The total number of users who have rated the game.                                             |
| url           | character | The URL to the game’s page on BoardGameGeek, providing more details and user reviews.          |
| thumbnail     | character | A URL to a thumbnail image of the game, often showing its box art or key visuals.              |

#### `details.csv`

| variable                | class     | description                                                                                     |
|:------------------|:------------------|:---------------------------------|
| num                     | double    | A unique identifier for each game within the dataset.                                           |
| id                      | double    | The BoardGameGeek ID for the game, used for linking to other datasets or resources.             |
| primary                 | character | The primary or most common name of the game, often used as its official title.                  |
| description             | character | A detailed textual description of the game, including its theme, mechanics, and objectives.     |
| yearpublished           | double    | The year the game was officially published or released to the public.                           |
| minplayers              | double    | The minimum number of players required to play the game.                                        |
| maxplayers              | double    | The maximum number of players that the game can accommodate.                                    |
| playingtime             | double    | The estimated total time, in minutes, needed to complete a game session.                        |
| minplaytime             | double    | The shortest estimated time required to complete a game session.                                |
| maxplaytime             | double    | The longest estimated time required to complete a game session.                                 |
| minage                  | double    | The minimum recommended age for players, often based on game complexity and content.            |
| boardgamecategory       | character | A categorical label describing the theme or genre of the game, such as "Strategy" or "Fantasy". |
| boardgamemechanic       | character | A textual description of the gameplay mechanics, such as "Deck Building" or "Dice Rolling".     |
| boardgamefamily         | character | A classification grouping games into families, often based on shared themes or settings.        |
| boardgameexpansion      | character | A list of expansions or add-ons related to the game, if applicable.                             |
| boardgameimplementation | character | A description of specific implementations or adaptations of the game.                           |
| boardgamedesigner       | character | The name(s) of the game’s designer(s), who created its mechanics and concept.                   |
| boardgameartist         | character | The name(s) of the artist(s) who contributed to the visual design of the game.                  |
| boardgamepublisher      | character | The name of the publishing company that released the game.                                      |
| owned                   | double    | The number of users who have marked the game as owned in their BoardGameGeek profile.           |
| trading                 | double    | The number of users who have listed the game as available for trade.                            |
| wanting                 | double    | The number of users who have expressed a desire to own the game.                                |
| wishing                 | double    | The number of users who have added the game to their wish list.                                 |

## Data Import and Cleaning

Import data for board game ratings and details.

```{r,warning=FALSE,message=FALSE}
# Load libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(janitor)
library(tidytext)
library(tidyverse) 
library(knitr)
library(stringr)

ratings <- readr::read_csv('ratings.csv')
details <- readr::read_csv('details.csv')

#kable(head(ratings))
#kable(head(details))
```

Perform data cleaning: handle missing values, correct data types, and remove duplicates.

```{r}
details <- details %>%
  clean_names()
  #rename_with(~ gsub("^boardgame", "", .), starts_with("boardgame"))

colSums(is.na(details))

details <- details %>%
  mutate(across(c(yearpublished, minplayers, maxplayers, minage), as.integer))

details <- details %>%
  mutate(across(where(is.numeric), ~ replace_na(., 0))) %>%
  mutate(across(where(is.character), ~ replace_na(., "Unknown")))

merged_data <- details %>%
  left_join(ratings %>% clean_names(), by = "num")
```

## Data Exploration

Let’s start by taking a broad look at the landscape of board games over time by examining the top 10 most popular games:

```{r}
top_10_games <- ratings %>%
  filter(!is.na(year) & rank <= 10) %>%
  arrange(rank)

ggplot(top_10_games, aes(x = year, y = rank, label = name)) +
  geom_point(color = "blue") +
  geom_text(size = 3, vjust = -0.5) +
  labs(title = "Top 10 Board Games Over Time", 
       x = "Year", 
       y = "Global Rank")
```

#### **What is the most common board game category?**

Rather than a comprehensive analysis, lets establish a focus and first identify the most common category of board games.

```{r}

category_counts <- merged_data %>%
  filter(boardgamecategory != "Unknown") %>%
  separate_rows(boardgamecategory, sep = ",") %>%
  count(boardgamecategory, sort = TRUE)

ggplot(category_counts[1:10, ], aes(x = reorder(boardgamecategory, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Board Game Categories", 
       x = "Category", 
       y = "Frequency") +
  theme_minimal()

most_common_category <- details %>%
  separate_rows(boardgamecategory, sep = ", ") %>%
  group_by(boardgamecategory) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1)
```

From this, we see that Card Games are by far the most popular board games when compared to other game modes. However, this includes a lot of old basic card games that are no longer played much anymore. Lets see how times have changed to match more recent times.

#### **How do ratings for card games change over time (1990 to present)?**

To analyze how the average ratings for card games have evolved over time, we limit the data set to games published from 1990 onward.

```{r}

card_games <- details %>%
  filter(
    grepl("Card Game", boardgamecategory) & 
    yearpublished >= 1990
  )

card_games_data <- card_games %>%
  left_join(ratings %>% clean_names(), by = "num")

card_games_trends <- card_games_data %>%
  filter(!is.na(average)) %>%
  group_by(yearpublished) %>%
  summarize(avg_rating = mean(average, na.rm = TRUE))

ggplot(card_games_trends, aes(x = yearpublished, y = avg_rating)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(color = "darkblue", size = 2) +
  labs(
    title = "Average Ratings for Card Games (1990 to Present)",
    x = "Year Published",
    y = "Average Rating"
  )

```

We see in recent years, there has been a significant decline in the popularity of card games. This downward trend suggests a shift in entertainment preferences, as card games seem to be losing their appeal compared to other forms of games. Despite the decrease in overall popularity, certain game mechanics still stand out in terms of user ratings and engagement. Let's take a closer look at the mechanics that define high-rated card games, as these reflect the qualities that continue to attract dedicated players. Analyze mechanics for card games:

```{r}

card_mechanics <- card_games_data %>%
  separate_rows(boardgamemechanic, sep = ", ") %>%
  group_by(boardgamemechanic) %>%
  summarize(
    avg_rating = mean(average, na.rm = TRUE),
    count = n()
  ) %>%
  arrange(desc(avg_rating))
```

Filter mechanics with sufficient representation and show them

```{r}

# Filter mechanics with sufficient representation
top_mechanics <- card_mechanics %>%
  filter(count >= 5) %>%
  arrange(desc(avg_rating))

# View top mechanics
kable(head(top_mechanics))
```

Meanwhile, in highly-rated mechanics across all board games, mechanics that emphasize **resource management** and **economic systems** are prevalent, such as:

-   Automatic Resource Growth (7.51) and Increase Value of Unchosen Resources (7.45), which reward strategic planning over time.

```{=html}
<!-- -->
```
-   *Bribery* (7.41) and *Turn Order: Pass Order* (7.38), which incorporate negotiation and decision-making under constraints.

```{r}
card_game_descriptions <- card_games_data %>%
  filter(description != "Unknown") %>%
  select(description)

word_frequency <- card_game_descriptions %>%
  unnest_tokens(word, description) %>%
  filter(!word %in% stop_words$word) %>% 
  count(word, sort = TRUE)

ggplot(word_frequency %>% slice_head(n = 20), aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(
    title = "Most Common Words in Card Game Descriptions",
    x = "Word",
    y = "Frequency"
  )

```

It appears that the most common word is simply "10," which doesn't provide much insight. Let's focus on identifying meaningful bigrams and filter out any instances of repeated numbers from the list of common words.

```{r}

card_game_bigrams <- card_game_descriptions %>%
  unnest_tokens(bigram, description, token = "ngrams", n = 2) 

separated_bigrams <- card_game_bigrams %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  filter(
    !word1 %in% stop_words$word, 
    !word2 %in% stop_words$word,
    !str_detect(word1, "^\\d+$"), 
    !str_detect(word2, "^\\d+$") 
  )

bigram_frequency <- separated_bigrams %>%
  unite(bigram, word1, word2, sep = " ") %>%
  count(bigram, sort = TRUE)

# Visualize the top 20 bigrams as a bar plot
ggplot(bigram_frequency %>% slice_head(n = 20), aes(x = reorder(bigram, n), y = n)) +
  geom_bar(stat = "identity", fill = "coral") +
  coord_flip() +
  labs(
    title = "Most Common Bigrams in Card Game Descriptions (Excluding Numbers)",
    x = "Bigram",
    y = "Frequency"
  ) 

```

## Conclusion

The BoardGameGeek dataset reveals several key insights into the trends, preferences, and ratings of board games over time. We observed a notable decline in the popularity of card games, especially in recent years, suggesting a shift in gaming preferences. The most common board game categories are dominated by card games, although this includes older games that are less relevant today. Additionally, mechanics that focus on strategic elements, resource management, and player interactions tend to receive higher ratings across both card games and board games in general. The bigram analysis of game descriptions also highlighted the prevalence of through instructions for strategic themes.
